### 2017-12-09 David Project/Paper/paper_mnras/paper.tex

% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
% \documentclass[a4paper,fleqn,usenatbib]{mnras}
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
%\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{subcaption}
\usepackage{graphicx}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
% macros. please check here before defining something new.
\input{../paper_resources/macros/code}
\input{../paper_resources/macros/derivatives}
\input{../paper_resources/macros/nuclides}
\input{../paper_resources/macros/concepts}
\input{../paper_resources/macros/units}
\input{../paper_resources/macros/vectors}
\input{../paper_resources/macros/formatting}
\input{../paper_resources/macros/symbols}

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
% misc. abbreviations
\newcommand{\paperone}{Paper~I} % the first paper

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% NuGrid/MESA way
%%
% see https://en.wikibooks.org/wiki/LaTeX/Colors for selection of
% colors, such as Apricot Aquamarine BlueGreen BurntOrange
% CornflowerBlue Emerald Gray Lavender Maroon NavyBlue Orchid Plum Red
% RoyalBlue SeaGreen Tan Violet YellowOrange

% for comments
\newcommand{\shortcomment}[3]{\textcolor{#1}{[#2: #3]}}
% for the Ready-To-Read sign off
\newcommand{\rtr}[1]{\shortcomment{red}{RTR}{#1}}
% for freezing a section
\newcommand{\sectionisfrozen}[1]{\textcolor{BlueViolet}{\textsf{\bfseries[section is frozen: send comments to #1]}}}
\newcommand{\sectionisdone}{\textcolor{Green}{\textsf{\bfseries[section is done]}}}

% if you want to embed comments in the text, clone the following with
% a color and your initials...

% for author's comments; choose your own color and replace 'fh' and
% 'FH' with your name and initials
\newcommand{\fhcom}[1]{\shortcomment{PineGreen}{FH}{#1}}

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Short title, max. 45 characters]{Ok Then, here it is}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. Stephens]{
David Stephens}

% These dates will be filled out by the publisher
%\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
%\pubyear{2015}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
During the asymptotic giant branch (AGB) phase of stellar evolution for a 2\Msun~star, the periodic thermal pulses have temperatures as high as 2.9\ee{8} \K. The \neon[22]($\alpha,n$)\magnesium[25]
After the formation of the \carbon[13] pocket

This is a simple template for authors to write new MNRAS papers.
The abstract should briefly describe the aims, methods, and main results of the paper.
It should be a single paragraph not more than 250 words (200 words for Letters).
No references should appear in the abstract.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
AGB -- Diffusion Coefficient -- Pre-solar Grains
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

* Why problem is interesting*

- Zr ratios in pre solar grains from AGB stars show isotopic signatures of material that left the star. 
- This is from envelope and Zr is produced in 13C pocket, then readjusted in the thermal pulse
- Tracer of conditions that were in the star at that time

* what have other people done *

- battino tried varying f and how diffusion happens for effects (couldnt get ratios seen)
- barzyk for xe128 ratio

* My probelm, the gap! *

- Try a different functionality of diffusion coefficient, hydro leads to this form
- It contains lower diffusion at base of convection zone, less 22Ne reaction, less Zr96

This is a simple template for authors to write new MNRAS papers.
See \texttt{mnras\_sample.tex} for a more complex example, and \texttt{mnras\_guide.tex}
for a full user guide.

All papers should start with an Introduction section, which sets the work
in context, cites relevant earlier studies in the field by \citet{1996ApJ...456..902R},
and describes the problem the authors aim to solve \citep[e.g.][]{cyburt:10}.

\section{Methods and Models}

Normally the next section describes the techniques the authors used.
It is frequently split into subsections, such as Section~\ref{sec:maths} below.

\subsection{\MESA Models}
\label{sec:mesa_models} % used for referring to this section from elsewhere

Only a 2\Msun, $Z=0.02$, stellar evolution model was usedd. This was computed using the \MESA~\citep{mesa} stellar code and taken from the \nugrid NuGrid set1ext, set1.2 models \citep{models}. These models were evolved from the pre-main sequence to a white dwarf. For this work, a singular thermal pulse, the 13th?, was analyzed and a Kippenhahn diagram of this particular thermal pulse can be seen in \Figure{kipp}.

For these models, the mixing lengths theory \citep{cox}, MLT, is used for the convection zones with a mixing length, $\alphaMLT=1.73$. Overshoot is implemented in \MESA with the formula from \citet{overshoot} and \citet{freytag}:

\begin{equation}
D = D_{0} exp^{-2z/f H_{p0}}
\label{eq:overshoot}
\end{equation}

Where D$_{0}$ and H$_{p0}$ are taken at the convective boundary.

\subsection{\zirconium[95] and \iodine[128] Branching}
\label{sec:branching}

\subsection{Diffusion Coefficient Modifications}
\label{sec:diffusion}

\subsection{\mppnp~Post Processing}
\label{sec:mppnp}

\subsection{Neutron Density and Temperature}
\label{sec:neutron}

\section{Results}
Simple mathematics can be inserted into the flow of the text e.g. $2\times3=6$
or $v=220$\,km\,s$^{-1}$, but more complicated expressions should be entered
as a numbered equation:

\begin{equation}
    x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}.
	\label{eq:quadratic}
\end{equation}

Refer back to them as e.g. equation~(\ref{eq:quadratic}).

% Kippenhahn figure
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/2M_Kippenhahn.png}
  \caption{Within the He intershell region, the \carbon[13] pocket forms and isotopic ratios are set by the \spr~(\Fig{zr_ratio13C}). It is then mixed and diluted in the He-flash pulse driven convection zone where temperatures get high enough to activate \neon[22]($\alpha,n$)\magnesium[25]. This shifts the isotopic ratios to those shown in \Figure{zr_ratioDil}} 
  \lFig{kipp}
\end{figure}

% c13 pocket zr ratio
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/C13_Zr.png}
  \caption{Within the \carbon[13] pocket there is a significant amount of \zirconium[94] produced but almost no \zirconium[96]. This is due to the low neutron densities not opening the \zirconium[95] branch.} 
  \lFig{zr_ratio13C}
\end{figure}

% Zr ratio after pulse
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/Pulse_Zr.png}
  \caption{After the \carbon[13] pocket is mixed into the He-flash pulse driven convection zone, the temperatures get high enough for the \neon[22]($\alpha,n$)\magnesium[25] reaction. This provides high enough neutron densities to open the \zirconium[95] branch and boost the \zirconium[96].} 
  \lFig{zr_ratioDil}
\end{figure}

% Xe plot
\begin{figure}
	\includegraphics[width=\columnwidth]{figs/xe.png}
    \caption{Another figure.}
    \lFig{mass-O16}
\end{figure}

% Zr 95 branch
\begin{figure}
	\includegraphics[width=\columnwidth]{figs/Zr95_branch.png}
    \caption{Another figure.}
    \lFig{zr95}
\end{figure}

% Zr 94 plot
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/Zr94.png}
  \caption{This is the Zr94 mil plot \citet{zr}.} 
  \lFig{zr94}
\end{figure}

% diffusion plot
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/Diffusion_compare.png}
  \caption{This is a comparison between the diffusion coefficients} 
  \lFig{diffusion}
\end{figure}

% time scale plot
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/Time_scale.png}
  \caption{This is the time scale plot} 
  \lFig{time_scale}
\end{figure}

% neutron denisty time plot
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/Neutron_Density_Time.png}
  \caption{This is the neutron density as a function of time} 
  \lFig{density}
\end{figure}

% sub time step neutron plots
\begin{figure}
\centering
\begin{subfigure}[b]{1\columnwidth}
   \includegraphics[width=1\columnwidth]{figs/T_Neutron_sub.png}
   \lFig{neutron_sub}
\end{subfigure}

\begin{subfigure}[b]{1\columnwidth}
   \includegraphics[width=1\columnwidth]{figs/T_Neutron_Sub_Closeup}
   \lFig{neutron_sub_close}
\end{subfigure}
\caption{Testing}
\end{figure}

Another results is also good \Fig{mass-O16}.
% Example table
\begin{table}
	\centering
	\caption{This is an example table. Captions appear above each table.
	Remember to define the quantities, symbols and units used.}
	\lTab{example_table}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		A & B & C & D\\
		\hline
		1 & 2 & 3 & 4\\
		2 & 4 & 6 & 8\\
		3 & 5 & 7 & 9\\
		\hline
	\end{tabular}
\end{table}

\section{Conclusions}

The last numbered section should briefly summarise what has been done, and describe
the final conclusions which the authors draw from their work.

\section*{Acknowledgements}

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{paper.bib} % if your bibtex file is
                                             % called example.bib

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Some extra material}

If you want to present additional material which would interrupt the
flow of the main paper, it can be placed in an Appendix which appears
after the list of references.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex